Script to make a map and add points
=========================================

Load libraries

```{r setup}
#sessionInfo()

library(raster)
library(rgdal) # NOTE: have to have [gdal](http://www.gdal.org/) and [PROJ.4](http://trac.osgeo.org/proj/) install on PATH
library(dismo)
library(maptools)
library(maps)
library(ggplot2)
```


Get climate data. Use highest scale resolution (10deg) for faster download at large scale. Other options are 5, 2.5 and 0.5 (which requires downloading individual tiles)

```{r loadclimdata}
w <- getData('worldclim', var='bio', res=10)
names(w)
# Plot first bioclim dimension (MAT) to check download worked
plot(w, 1)
```


Load file with location data. Longitude column ('lon') should precede latitude ('lat')

```{r loadsitedata}
d <- read.csv("Aphaeno2013_sampling_locations.csv")
head(d)
str(d)
```

Extract bioclimatic variables from worldclim data. Note bioclim temperature data is multiplied by 10 (no decimal), so transform by dividing by 10.

```{r extract}
# Extract climate values for site lon/lat from bioclim data
dbio1 <- extract(w, d[,c("lon","lat")])

dbio1 <- cbind(d, dbio1[,1])
colnames(dbio1)[5] <- "MATx10"
dbio1$MAT <- dbio1$MATx10/10
head(dbio1)

# Write to file
write.table(dbio1, file="Aphaen2013_sites_MAT.txt", quote=FALSE, sep="\t", row.names=FALSE)
```


Plot NorthEastern US and add points for sampling sites. Note MAT is still multiplied by 10.

```{r plot}
plot(w, 1, xlim=c(-90, -60), ylim=c(25,50), axes=TRUE)
points(dbio1$lon, dbio1$lat, col="black", pch=20, cex=0.75)
```


Create same plot using ggplot2, with help from [stackoveflow](http://stackoverflow.com/questions/9422167/how-do-i-plot-a-single-point-on-a-world-map-using-ggplot2)

TODO: need to add MAT to plot....

```{r ggplot}
eUSA <- map_data('state', region=c('florida', 'south carolina', 'north carolina', 'georgia', 'virginia', 'west virginia', 'maryland', 'delaware', 'new jersey', 'rhode island', 'new york', 'connecticut', 'massachusetts', 'pennyslvania', 'vermont', 'new hampshire', 'maine', 'alabama', 'tennessee', 'kentucky', 'ohio'))

p <- ggplot(legend=FALSE) +
geom_path( data=eUSA, aes(x=long, y=lat,group=group)) +
theme(panel.background = element_blank()) +
theme(panel.grid.major = element_blank()) +
theme(panel.grid.minor = element_blank()) +
theme(axis.text.x = element_blank(),axis.text.y = element_blank()) +
theme(axis.ticks = element_blank()) +
xlab("") + ylab("")

# Sites to add to plot:
sites <- d[,c("lon", "lat")]
p <- p + geom_point(data=sites, aes(lon, lat), colour="green", size=4)
p
```

Hmmm...looks like points are out of place...probably due to Google Earth giving DD'MM'SS and R plots expecting DD.DDDD

Would be cool to make animated plot showing changes in temperature over past 100 years using something like [this](http://r-nold.blogspot.com/2012/08/provincial-monthly-rainfall-of.html)


