Script to make a map and add points
=========================================

Load libraries

```{r setup}
library(raster)
library(dismo)
library(maptools)
```


Get climate data. Use highest scale resolution (10deg) for faster download at large scale. Other options are 5, 2.5 and 0.5 (which requires downloading individual tiles)

```{r loadclimdata}
w <- getData('worldclim', var='bio', res=10)
names(w)
# Plot first bioclim dimension (MAT) to check download worked
plot(w, 1)
```


Load file with location data. Longitude column ('lon') should precede latitude ('lat')

```{r loadsitedata}
d <- read.csv("Aphaeno2013_sampling_locations.csv")
head(d)
str(d)
```

Extract bioclimatic variables from worldclim data. Note bioclim temperature data is multiplied by 10 (no decimal), so transform by dividing by 10.

```{r extract}
# Extract climate values for site lon/lat from bioclim data
dbio1 <- extract(w, d[,c("lon","lat")])

dbio1 <- cbind(d, dbio1[,1])
colnames(dbio1)[5] <- "MATx10"
dbio1$MAT <- dbio1$MATx10/10
head(dbio1)

# Write to file
write.table(dbio1, file="Aphaen2013_sites_MAT.txt", quote=FALSE, sep="\t", row.names=FALSE)
```


Plot NorthEastern US and add points for sampling sites. Note MAT is still multiplied by 10.

```{r plot}
plot(w, 1, xlim=c(-90, -60), ylim=c(25,50), axes=TRUE)
points(dbio1$lon, dbio1$lat, col="black", pch=20, cex=0.75)
```

